---
title: "Running R, Python, and SQL in Rmarkdown"
subtitle: 'Using penguins dataset'
author: "Lucky Mehra"
date: "2/26/2021"
output:
  rmdformats::robobook:
    code_folding: show
    self_contained: true
    thumbnails: false
    lightbox: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

# set up for SQL
library(DBI)
library(palmerpenguins)

# connection
con <- dbConnect(RSQLite::SQLite(), dbname = ":memory:")

# write tables in database
dbWriteTable(con, "penguins", penguins)
```

## Subset dataset

Select columns `species`, `island`, `sex`, and `bill_length_mm`.

And filter rows where `body_mass_g` between 3000 and 4000.

### R

```{r}
library(tidyverse)
penguins %>% 
  dplyr::filter(body_mass_g <= 4000 & body_mass_g >= 3000) %>% 
  dplyr::select(species, island, sex, bill_length_mm)
```

### Python

```{python}
# import modules
import pandas as pd

# access penguins data from r chunks
penguins_py = r.penguins

# apply the filtering and selecting
(penguins_py
  .query('(body_mass_g <= 4000) & (body_mass_g >= 3000)') # equivalent to dplyr::filter
  .filter(['species', 'island', 'sex', 'bill_length_mm'])) # equivalent to dplyr::select
```

### SQL

```{sql, connection=con}
SELECT species,
        island, 
        sex, 
        bill_length_mm
FROM penguins
WHERE body_mass_g BETWEEN 3000 AND 4000
```

## Create new columns

### R

```{r}
penguins %>% 
  mutate(ratio_length_depth = bill_length_mm / bill_depth_mm) %>% 
  dplyr::select(species, island, bill_length_mm,
                bill_depth_mm, ratio_length_depth)
```
### Python

```{python}
# create new column
penguins_py['ratio_length_depth'] = penguins_py.bill_length_mm / penguins_py.bill_depth_mm

# print desired columns
(penguins_py
  .filter(['species', 'island', 'bill_length_mm',
  'bill_depth_mm', 'ratio_length_depth']))
```

### SQL

```{sql connection=con}
SELECT species,
        island,
        bill_length_mm,
        bill_depth_mm,
        bill_length_mm / bill_depth_mm AS ratio_length_depth --create new column
FROM penguins
```

## Do 'group by'

Calculate average, minimum, and maximum of `bill_length_mm` by `species`. Then, arrange/sort/order
the species from highest mean bill length to lowest mean bill length i.e. the species with the
highest mean bill length should be the first row.

### R

```{r}
penguins %>% 
  group_by(species) %>% 
  summarize(mean_bill_length = mean(bill_length_mm, na.rm = TRUE),
            min_bill_length  = min(bill_length_mm, na.rm = TRUE),
            max_bill_length = max(bill_length_mm, na.rm = TRUE)) %>% 
  arrange(mean_bill_length %>% desc())
```

### Python

```{python}
# import pandas as pd
(penguins_py
  .groupby('species')
  .bill_length_mm
  .agg(['mean', 'min', 'max'])
  .sort_values('mean', ascending=False))
```

### SQL

```{sql connection=con}
SELECT species,
        AVG(bill_length_mm) AS mean_bill_length,
        MIN(bill_length_mm) AS min_bill_length,
        MAX(bill_length_mm) AS max_bill_length
FROM penguins
GROUP BY species
ORDER BY mean_bill_length DESC
```










